---
title: "Tic Tac Toe"
output: github_notebook
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(glue)
```


## Notes
- Taking in input from the prompt is not a functionality I have used before in R, but it definately exists (see eg packages built to provide in-IDE tutorials from an R package such as `swirl`)

### R features
- https://www.datamentor.io/r-programming/examples/user-input/
- https://stackoverflow.com/questions/33425727/calling-user-inputs-prompted-through-one-r-function-in-a-second-r-function 

- Existing R package (haven't looked at the code) https://github.com/kota7/tictactoe 
- Implementation that includes a way to code  grid to plot: https://www.robert-hickman.eu/post/r-inforcement_learning_one/ 

### Tic tac toe and deadling with user input in other languages
- https://towardsdatascience.com/tic-tac-toe-creating-unbeatable-ai-with-minimax-algorithm-8af9e52c1e7d
- https://stackoverflow.com/questions/23294658/asking-the-user-for-input-until-they-give-a-valid-response


## Sample function 
```{r}
# Sample function 
initial_prompt <- function()  {
  print(glue("
                                                                     
             Welcome to tic-tac-toe!
             
             This is a game for two players. The goal is to get three x's or o's in a row on the board!

             You can take turns entering your move in this window. 
             There are a total of 9 positions on the board, as show below: 
                
            +---+---+---+
            | 1 | 2 | 3 |
            +---+---+---+
            | 4 | 5 | 6 |
            +---+---+---+
            | 7 | 8 | 9 |
            +---+---+---+"
                         
             ))
  
  
  print(glue("Let's start the game! The board is blank, and ready for the first player: 
             
             +--+--+--+
             |  |  |  |
             +--+--+--+
             |  |  |  |
             +--+--+--+
             |  |  |  |
             +--+--+--+
             
             "))
  
  n <- (readline("Player 1, what's your move? Please enter a number from 1-9: "))
  n <- as.numeric(n)
  
  list(n=n)
}

tic_tac_toe <- function ()  {
  
  # Initialize to store moves from each player
  player_1_moves <- c()
  player_2_moves <- c()
  
  inputs <- initial_prompt()
  
  player_1_moves <- c(player_1_moves, inputs$n)
  
  print(glue("Moves from player 1 so far:   {player_1_moves}"))
}
```

## Steps and desired feautures

- Two humans play against each other
- A message when the function is first called: eg  "Welcome to tic tac toe! This is a game for two players. The goal is to get three x's or o's in a row!

You can take turns entering your move in this window. There are a total of 9 positions on the board, as show below: 
"
- Optional - flip a coin to decide who goes first

- Initialize sets to store moves from each player

- A new message that shows the state of the board and promps one player to enter their move (from 1 to 9)

- The number the first player enters is added to their set of moves

- State of board is updated, and the second player is prompted to enter their move

- Loop three steps above, evaluating win conditions at each iteration
    - There are 8 possible configurations for each player to win, which are easy to list as a set (order doesn't matter)


