---
title: "Tic Tac Toe"
output: github_notebook
editor_options: 
  chunk_output_type: console
---

```{r setup, message= F}
library(glue)  # for helpful defaults for print statements in R
library(purrr) # For functions to map over a vector (specifically `map_int()` below)
library(magrittr) # For 'pipe' operator ( %>% ) to specify operations from left to right w/o nested function calls
```

## Functions

### Helper functions
```{r}

initial_prompt <- function()  {
  
      
            #Initialize
             player_1_moves <- c()
             player_2_moves <- c()
  
  print(glue("
                                                                     
             Welcome to tic-tac-toe!
             
             This is a game for two players. The goal is to get three x's or o's in a row on the board 
             (in any order - horizontal, vertical, or diagonal)

             You can take turns entering your move at this prompt. 
             There are a total of 9 positions on the board, each indicated with a number, as shown below:
             
             ")) 
        
                
             plot_board(player_1_moves, player_2_moves)
                        
  print(glue("
             
                          Player 1 will go first. Your plays will use the 'x' symbol on the board. Player 2 is 'o'.
             
             "))
                     
  
  n <- (readline("Player 1, what's your first move? Please enter a number from 1-9: "))
  n <- as.numeric(n)
  
  list(n=n)
}

 player2_initial_prompt <- function(){
    n <- (readline("Player 2, it's time for your first move with 'o', please enter a number from 1-9: "))
    n <- as.numeric(n)
  
  list(n=n)
 }
 
midgame_prompt <- function(player_number){
    n <- (readline(glue("Player {player_number}, your turn. Please enter a number from 1-9: ")))
    n <- as.numeric(n)
  
  list(n=n)
 }
 
```

```{r}
# Plotting the board - shows all digits if player_moves is NULL 
# Initially adapted from example that used a v. different numbering scheme https://www.robert-hickman.eu/post/r-inforcement_learning_one/
plot_board <- function(player_1_moves, player_2_moves ) {
  
  board_numbers <- c(1:9)
  
   for (i in board_numbers)
  {
    if (i %in% player_1_moves)  board_numbers[i] <- "x"
    if (i %in% player_2_moves)  board_numbers[i] <- "o"
    }
  
  board_elements <- glue_collapse(board_numbers, sep = " | ")
 
  board_lines <- gsub("(. \\| . \\| . )\\|( . \\| . \\| . )\\|( . \\| . \\| .)", 
                      "\n \\1\n-----------\n\\2\n-----------\n\\3",
                      board_elements
                      )
  return(writeLines(board_lines))
}
```

```{r}

play_again_prompt <- function() {
    c <- (readline(glue("Would you like to play again? Please enter Y for 'Yes' or N for 'No': ")))
    c <- as.character(c)
  
  list(c = c)
}  
```

```{r}

tic_tac_toe <- function ()  {
  
    # Initialize empty vectors to store moves from each player
    player_1_moves <- c()
    player_2_moves <- c()
  

   #### First move by player 1 ####
   inputs <- initial_prompt() # Trigger text input prompt
  
   p1_next_move <- inputs$n # Extract value from text input prompt
   player_1_moves <- append(player_1_moves, p1_next_move)
   
   plot_board(player_1_moves, player_2_moves) 
   
   
   #### First move by player 2 ####
   inputs2 <- player2_initial_prompt()
   
   p2_next_move <- inputs2$n
   player_2_moves <- append(player_2_moves, p2_next_move)
   
   plot_board(player_1_moves, player_2_moves)
   
   #### Initialize conditions ####
   results <- list(continue_play = T, winner = NA)
   
   #### Function to test for win conditions #### 
   evaluate_win_conditions <-function(player_1_moves, player_2_moves) {

      #Enumerate winning combinations of positions (as sets where elements can occur in any order)
      list_of_winning_moves <- list(c(1,2,3), c(4,5,6), c(7,8,9), c(1,4,7), c(2,5,8), c(3,6,9), c(1,5,9), c(3,5,7))
  
      # Map return a list, map_int simplifies the output to a vector of integers
      p1_results <- map_int(list_of_winning_moves, function(x) is.element(unlist(x), player_1_moves) %>% sum() )  
      p2_results <- map_int(list_of_winning_moves, function(x) is.element(unlist(x), player_2_moves) %>% sum() ) 

      if (max(p1_results) == 3 ) {
          results$continue_play <- F
          results$winner <- "Player 1" 
        } else if (max(p2_results) == 3 ){ 
          results$continue_play <- F
          results$winner <- "Player 2" }
        
       results <- list(continue_play = results$continue_play, winner =  results$winner)
      }
  
  #### Function to check if input is numeric and has not been called already ####
   
    ## Note: Is inside tic_tac_toe function because otherwise `player_1_moves` is
    ## not found when this function called from inside the larger function
  test_numeric_move <- function(next_move) {
  
    # First check if input is numeric 
       if (is.na(next_move) == F) {   # Input is corerced to numeric in midgame_prompt, so NA if not a number
        error1 <- F
      } else { 
          error1 <- T
          print(glue("Oops! Looks like you entered something that isn't a number, so your move isn't on the board yet.
                        Please try again and enter a number between 1 and 9"))
      }
    
    # Second, check if input is within range 1 - 9, for inputs that are numeric (not NA)
      if (next_move %in% c(1,2,3,4,5,6,7,8,9)) { 
        error2 <- F
      } else if (is.na(next_move)){ error2 <- T # no additional message, given message above
      } else {
          error2 <- T
          print(glue("Sorry, you entered '{next_move}', which is not a position on the board.
                        Please try again and enter a number between 1 and 9 that hasn't been played yet"))
            }
      
    # Third check if the numeric value has already been played elsewhere on the board (numbers can't overlap)
      if (next_move %in% player_1_moves | next_move %in% player_2_moves) {
          error3 <- T
          print(glue("Hmm, you entered '{next_move}', but someone has already used that spot on the board
                        Please try again and enter a position on the board that is still avalible"))
      } else { 
        error3 <- F 
      }
      
      # If either condition failed, need to prompt player to re-enter
      need_reenter <- ifelse(error1 == T | error2 == T | error3 == T, T, F) # If either error is T, then need_reenter is T 

}
   
  #### While loop with next moves by both players, including input checking ####
   # Version w/ two while loops inside while loop for (not optimal) error handling
   
   while (results$continue_play == T ){
     
     # (Re)initialize for each round of play
     need_reenter_p1 <- T
     need_reenter_p2 <- T
     
        # Player 1
        while (need_reenter_p1 == T ){
    
          player1_inputs <- midgame_prompt(player_number = 1)
  
          # Check content of input, and re-prompt player from start of loop if needed
          need_reenter_p1 <- test_numeric_move(p1_next_move <- player1_inputs$n)
             if ( need_reenter_p1 == T ) {
               next
             }else if (need_reenter_p1 == F) {
               player_1_moves <- append(player_1_moves, p1_next_move)
             }
        } # end of player 1 while loop
        
          plot_board(player_1_moves, player_2_moves)
      
          results <- evaluate_win_conditions(player_1_moves, player_2_moves) # Check if either player has won yet
      
     if (results$continue_play == F) { break }
     else {
      
        # Player 2
        while (need_reenter_p2 == T ){
    
          player2_inputs <- midgame_prompt(player_number = 2)
  
          # Check content of input, and re-prompt player from start of loop if needed
          need_reenter_p2 <- test_numeric_move(p2_next_move <- player2_inputs$n)
             if ( need_reenter_p2 == T ) {
               next
             }else if (need_reenter_p2 == F) {
               player_2_moves <- append(player_2_moves, p2_next_move)
             }
        } # end of player 2 while loop
        
          plot_board(player_1_moves, player_2_moves)
      
          results <- evaluate_win_conditions(player_1_moves, player_2_moves) # Check if either player has won yet
        
      } # end of else statement
       } # end of umbrella while loop
    
    #### End of game  - check if want to play again ####
   
        print(glue("Congrats, {results$winner}, you won! "))
        
         yes_no <- play_again_prompt()
         yes_no_choice <- yes_no$c
         
         if(yes_no_choice %in% c("N","n")) {
           
           print(glue("Game over"))
          
           } else if(yes_no_choice %in% c("Y","y")) {
             
             tic_tac_toe()
             
           } else {print(glue("Sorry, '{yes_no_choice}' is not a valid input"))
             
                # Repeat choices above
                yes_no <- play_again_prompt()
                yes_no_choice <- yes_no$c
                
                if(yes_no_choice %in% c("N","n")) {
                  print(glue("Game over"))
                  } else if(yes_no_choice %in% c("Y","y")) {
                    tic_tac_toe()
                  } else {print(glue("Sorry, this doesn't seem to be working out. Game over."))}
                    
            } # End of else statement
             
        } # End of function
     
  
```

```{r}

# Will writing a for loop with i in some_sequence mean 'next' works for error handling to re-prompt first user? No

for (i in c(1,2)) {
  print(glue("i is : {i}")) # i is either player 1 or 2
  
  input <- midgame_prompt(player_number = i)
  
  need_reenter <- test_numeric_move(next_move <- input$n)
  
  if (need_reenter == T) {
    next
  }else if (need_reenter == F) {
             player_moves[[i]] <- append(player_moves[[i]], p1_next_move)
           }
}
```

```{r}

# Is it possible to write a for loop where there is only 1 entry in the sequence?
for (i in 1) {
  print(glue("i is : {i}")) # i is either player 1 or 2
  
  input <- midgame_prompt(player_number = i)
  
  need_reenter <- test_numeric_move(next_move <- input$n)
  
  if (need_reenter == T) {
    next
  }else if (need_reenter == F) {
             player_moves[[i]] <- append(player_moves[[i]], p1_next_move)
           }
  }

```


## Interim notes

- R error and warning handling is canonically through `tryCatch`, which I have not used before:  see https://advanced-r-solutions.rbind.io/conditions.html#signalling-conditions,  https://stackoverflow.com/questions/8093914/use-trycatch-skip-to-next-value-of-loop-upon-error , 

_ Lists in R *are * ordered, even though they can also be named - good writeup: https://stackoverflow.com/questions/2050790/how-to-correctly-use-lists-in-r 
- R base set operations are set up to work on vectors
- Trying to create a tibble from vectors like this does not work and only reads the first vector: `as_tibble(c(1,2,3), c(4,5,6), c(7,8,9), c(1,4,7), c(2,5,8), c(3,6,9), c(1,5,9), c(3,5,7))`
- sapply is a simplified version of lapply output (lapply returns a list same length as original element)

- Can use combinat::permn to generate all permutations of a vector
- Can use `purrr::map2()` to operate on 2 things eg "(data = DATA FOR ONE SPECIES, n = SAMPLE SIZE)", per https://jennybc.github.io/purrr-tutorial/ls12_different-sized-samples.html

- Taking in input from the prompt is not a functionality I had used before in R, but it definately exists (see eg packages built to provide in-IDE tutorials from an R package such as `swirl`)
  - https://www.datamentor.io/r-programming/examples/user-input/
  - https://stackoverflow.com/questions/33425727/calling-user-inputs-prompted-through-one-r-function-in-a-second-r-function 

- Existing R package (haven't looked at the code) https://github.com/kota7/tictactoe 

## Steps and desired features

- Two humans play against each other
- A message when the function is first called: eg  "Welcome to tic tac toe! This is a game for two players. The goal is to get three x's or o's in a row!

You can take turns entering your move in this window. There are a total of 9 positions on the board, as show below: 
" - Done
- Optional - flip a coin to decide who goes first

- Initialize sets to store moves from each player - Done

- A new message that shows the state of the board and promps one player to enter their move (from 1 to 9) - Done

- The number the first player enters is added to their set of moves - Done

- State of board is updated, and the second player is prompted to enter their move - Done

- Loop three steps above, evaluating win conditions at each iteration
    - There are 8 possible configurations for each player to win, which are easy to list as a set (order doesn't matter)

### Tic tac toe and dealing with user input in other languages
- https://towardsdatascience.com/tic-tac-toe-creating-unbeatable-ai-with-minimax-algorithm-8af9e52c1e7d
- https://stackoverflow.com/questions/23294658/asking-the-user-for-input-until-they-give-a-valid-response

