---
title: "Tic Tac Toe"
output: github_notebook
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(glue)
library(tidyverse)
```


## Notes
- Taking in input from the prompt is not a functionality I have used before in R, but it definately exists (see eg packages built to provide in-IDE tutorials from an R package such as `swirl`)

### R features
- https://www.datamentor.io/r-programming/examples/user-input/
- https://stackoverflow.com/questions/33425727/calling-user-inputs-prompted-through-one-r-function-in-a-second-r-function 

- Existing R package (haven't looked at the code) https://github.com/kota7/tictactoe 
- Implementation that includes a way to code  grid to plot: https://www.robert-hickman.eu/post/r-inforcement_learning_one/ 

### Tic tac toe and deadling with user input in other languages
- https://towardsdatascience.com/tic-tac-toe-creating-unbeatable-ai-with-minimax-algorithm-8af9e52c1e7d
- https://stackoverflow.com/questions/23294658/asking-the-user-for-input-until-they-give-a-valid-response


## Sample function 
```{r}
# Sample function 
initial_prompt <- function()  {
  print(glue("
                                                                     
             Welcome to tic-tac-toe!
             
             This is a game for two players. The goal is to get three x's or o's in a row on the board!

             You can take turns entering your move in this window. 
             There are a total of 9 positions on the board, as show below: 
                
            +---+---+---+
            | 1 | 2 | 3 |
            +---+---+---+
            | 4 | 5 | 6 |
            +---+---+---+
            | 7 | 8 | 9 |
            +---+---+---+"
                         
             ))
  
  
  print(glue("Let's start the game! The board is blank, and ready for the first player: 
             
             +--+--+--+
             |  |  |  |
             +--+--+--+
             |  |  |  |
             +--+--+--+
             |  |  |  |
             +--+--+--+
             
             "))
  
  n <- (readline("Player 1, what's your move? Please enter a number from 1-9: "))
  n <- as.numeric(n)
  
  list(n=n)
}

tic_tac_toe <- function ()  {
  
  #Enumerate winning combinations of positions
  winning_moves <- list(c(1,2,3), c(4,5,6), c(7,8,9))
  
  # Initialize to store moves from each player
  player_1_moves <- c()
  player_2_moves <- c()
  
  inputs <- initial_prompt()
  
  player_1_moves <- c(player_1_moves, inputs$n)
  
  print(glue("Moves from player 1 so far:   {player_1_moves}, type of player1_moves = {class(player_1_moves)}"))
}
```

```{r}
# Test tracking winning moves
# One straightforward  way to do this seems to be as a vector, with the character strings of all possible results
# Use `permn` to genertae permutations

vector_of_winning_moves <- c("123", "132", "312", "321", "231", "213")

test_vector <- c("123")

test_vector %in% vector_of_winning_moves



```


```{r}
# Notes on ways to use lists and sets instead that didn't make sense

# Defining as list of vectors
winning_moves <- list(c(1,2,3), c(4,5,6), c(7,8,9), c(1,4,7), c(2,5,8), c(3,6,9), c(1,5,9), c(3,5,7))


test_set_of_moves_1 <- list(c(3,1,2))

test_set_of_moves_2 <- c(4,5,6)

# Not as expected - c(4,5,6) is clearly an element in winning_moves
test_set_of_moves_2 %in% winning_moves
winning_moves %in% test_set_of_moves_2 

test_set_of_moves_1 %in% winning_moves
winning_moves %in% test_set_of_moves_1 


# I thought intersect would be the way to do this, but the outputs are blank data types
intersect(test_set_of_moves_1, winning_moves )
intersect(winning_moves, test_set_of_moves_1 )

intersect(test_set_of_moves_2, winning_moves )
intersect(winning_moves, test_set_of_moves_2 )
```

```{r}
# Alternate example not with numbers - why does this work with characters and not with numbers?
check_against <- c("a", "b", "c", "d")

"b" %in% check_against

c("a", "c") %in% check_against


# alternate example that also includes list - checking for element does not work
check_against_list <- list(c("a", "b", "c"), c("d", "e", "f"), c("g", "h", "i"))

# Check if "b" is in any of the lists - Works with list comprehension ("b" %in% check_against_list does not)
# https://stackoverflow.com/questions/53086053/how-to-check-if-a-list-contains-a-certain-element-in-r 

1 * sapply(check_against_list, function(x) "b" %in% x)  

# Check if c("a", "b", "c") is in any of the lists 
lapply(check_against_list, function(x) c("a", "b","c") %in% x)  
1 * sapply(check_against_list, function(x) c("a", "b","c") %in% x)  

# Check if the elements, in a different order, are in the list - also works
lapply(check_against_list, function(x) c("c", "a","b") %in% x)  

#Check if the elements, in a different order, are in the list with tidyverse syntax

map_lgl(check_against_list, `%in%`, x ="b") %>% 
    as.integer

map(check_against_list, `%in%`, x = c("b", "c", "a")) %>% unlist() %>% as.numeric() %>% sum()

map(check_against_list, `%in%`, x = c("b", "e", "g")) %>% unlist() %>% as.numeric() %>% sum()

```

_ Lists in R *are * ordered, even though they can also be named - good writeup: https://stackoverflow.com/questions/2050790/how-to-correctly-use-lists-in-r 
- R base set operations are set up to work on vectors
- Trying to create a tibble from vectors like this does not work and only reads the first vector: `as_tibble(c(1,2,3), c(4,5,6), c(7,8,9), c(1,4,7), c(2,5,8), c(3,6,9), c(1,5,9), c(3,5,7))`
- sapply is a simplified version of lapply output (lapply returns a list same length as original element)

- Can use combinat::permn to generate all permutations of a vector


```{r}
new_list_of_winning_moves <- check_against_list <- list(c("1", "2", "3"), 
                                                        c("1", "3", "2"), 
                                                        c("3", "1", "2"),
                                                        c( "3", "2", "1"),
                                                        c("2", "3", "1"),
                                                        c("2", "1", "3"))

test_vector <- c("1", "2", "3")

map(new_list_of_winning_moves, `%in%`, x =test_vector) 

non_matching_test_vector <- c("1", "4", "3")

map(new_list_of_winning_moves, `%in%`, x =non_matching_test_vector) 


# Try intersect
map(new_list_of_winning_moves, function(x) intersect(x, non_matching_test_vector)) 

map(new_list_of_winning_moves, function(x) intersect(x, test_vector)) 

# What if make input a list of lists? Does not work as expected

list_list_of_winning_moves <- list(list(c("1", "2", "3")), 
                                   list(c("1", "3", "2")), 
                                   list(c("3", "1", "2")),
                                   list(c("3", "2", "1")),
                                   list(c("2", "3", "1")),
                                   list(c("2", "1", "3")))

test_list <- c("1", "2", "3")

map(list_list_of_winning_moves, `%in%`, x =test_list) 

# What if make input a list of character vectors?

vector_of_winning_moves <- c("123", "132", "312", "321", "231", "213")

test_vector <- c("123")

test_vector %in% vector_of_winning_moves



```

## Steps and desired feautures

- Two humans play against each other
- A message when the function is first called: eg  "Welcome to tic tac toe! This is a game for two players. The goal is to get three x's or o's in a row!

You can take turns entering your move in this window. There are a total of 9 positions on the board, as show below: 
"
- Optional - flip a coin to decide who goes first

- Initialize sets to store moves from each player

- A new message that shows the state of the board and promps one player to enter their move (from 1 to 9)

- The number the first player enters is added to their set of moves

- State of board is updated, and the second player is prompted to enter their move

- Loop three steps above, evaluating win conditions at each iteration
    - There are 8 possible configurations for each player to win, which are easy to list as a set (order doesn't matter)


